---
title: "In Class Assignment Feb 28"
author: "Molly"
date: "2/28/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. 
```{r}
ckmnodes = read.csv(file = "ckm_nodes.csv")
dim(ckmnodes)
colnames(ckmnodes)
```

2. 16 doctors never prescribed tetracycline. 121 values are NA. See table below for counts at the end of each month. 
```{r}
table(ckmnodes$adoption_date)
names(which.max(table(ckmnodes$adoption_date)))
library('plyr')
count(ckmnodes, 'adoption_date')
```
3. 
```{r}
newvec = which(!is.na(ckmnodes$adoption_date))
length(newvec)
cleaned_nodes = ckmnodes[newvec,]
head(cleaned_nodes)
```
4. 
```{r}
adopters = function(month, not.yet = FALSE){
  if (not.yet == FALSE) {return (as.vector(cleaned_nodes$adoption_date ==  month))}
  else {return (as.vector(cleaned_nodes$adoption_date > month | cleaned_nodes$adoption_date == "Inf"))}
}
sum(adopters(2))
sum(adopters(month = 14, not.yet = TRUE))

## for a particular month, this function gives the number of doctos who began prescribing that month, unless not.yet is true, in which case it will return a vector of the doctors who began prescribing after that month or never 
```

5. 
```{r}
ckmnetwork = read.table(file = "/Users/mollymccarthy/Downloads/ckm_network (2).dat")
head(ckmnetwork)
dim(ckmnetwork)

newvec = which(!is.na(ckmnodes$adoption_date))
clean_network = ckmnetwork[newvec, newvec]
head(clean_network)
dim(clean_network)
```
6. 
```{r}
numberofcontacts = apply(clean_network, 1, sum)
is.vector(numberofcontacts)
numberofcontacts[41]
```
7a.  
```{r}
count_peer_pressure = function(doctor){
  length(which(cleaned_nodes$adoption_date[clean_network[,doctor]==1]<= cleaned_nodes$adoption_date[doctor]))
}

#this function finds all the contacts of doctor (clean_network[,doctor] ==1) and gives you their adoption date if that adoption date is less than or equal to the adoption date of doctor. It then gives the length of the resulting vector 
count_peer_pressure(37)  
```
  
```{r, eval = FALSE} 

##INCORRECT ATTEMPTS AT CREATING PEER PRESSURE FUNCTION
  contacts = which(clean_network[,doctor] == 1)  ##gives you a list of the contacts of the original doctor
  return(length(contacts[cleaned_nodes$adoption_date < month & cleaned_nodes$adoption_date == month]))#now we want to take these contacts and return all of the contacts of doctor for whom cleaned_nodes$adoption_date <= month
}

count_peer_pressure(37, 5)



for(i in 1:ncol(clean_network)){
  which(clean_network[,i] == 1)
}


## cleaned_nodes$adoption_date[clean_network[,1] == 1] gives you the month in which the contacts of doctor 1 began prescribing 
```
7b. 
```{r}
prop_peer_pressure = function(doctor){length(which(cleaned_nodes$adoption_date[clean_network[,doctor] == 1] <= cleaned_nodes$adoption_date[doctor])) / length(cleaned_nodes$adoption_date[clean_network[,doctor] ==1]) 
}
#takes the function from 7a and divides it by the total number of contactss
prop_peer_pressure(37)
prop_peer_pressure(102)
```
8a. 
```{r}

samemonthprop = function(month){
mean(apply(as.array(which(cleaned_nodes$adoption_date == month)), 1, prop_peer_pressure), na.rm = TRUE) 
}
##of all doctors who started prescribing in a particular month, what proportion of them had contacts who were already prescribing by that month 
samemonthprop(8)

latermonthsprop = function(month){
  mean(apply(as.array(which(cleaned_nodes$adoption_date > month |  cleaned_nodes$adoption_date == "Inf")), 1, prop_peer_pressure), na.rm = TRUE)
}
latermonthsprop(8)
#for all doctors who started prescribing after the month or never, what proportion of their contacts had already started prescribing 

averaging.peer.pressure = function(month){
  cbind(samemonthprop(month), latermonthsprop(month))
}

##puts two vectors together 
```

```{r, eval = FALSE}

##INCORRECT ATTEMPTS AT CREATING FUNCTION 
my_function = function(month, not.yet = FALSE){
  if(not.yet == FALSE) {return (mean(apply(as.array(which(cleaned_nodes$adoption_date == month)), 1, prop_peer_pressure), na.rm = TRUE))}
  else 
    {return (mean(apply(as.array(which(cleaned_nodes$adoption_date > month |  cleaned_nodes$adoption_date == "Inf")), 1, prop_peer_pressure), na.rm = TRUE))
  }
}
 ##if a doctor has begun prescribing by that month, then this function takes the mean proportion of contacts of that doctor who were prescribing on or before that month. If a doctor has not begun prescribing by that month, this function returns the average proportion of prescribers among that contact 

my_function(8)
```

```{r, eval = FALSE}
test = apply(as.array(which(cleaned_nodes$adoption_date ==8)), 1, prop_peer_pressure) ##all doctors who started prescribing in month 8, applies prop peer pressure function to all those doctors) 

mean(test) #finds the average proportion of prescribers among the contacts of the doctors who began prescribing in month 8

test2 = apply(as.array(which(cleaned_nodes$adoption_date > 8 | cleaned_nodes$adoption_date == "Inf")), 1, prop_peer_pressure) ##does same thing is other apply function but for doctors who prescribed after month 8 or never 

mean(test2, na.rm = TRUE)




doctors.in.month = which(cleaned_nodes$adoption_date == 5)

?apply

adopters = function(month, not.yet = FALSE){
  if (not.yet == FALSE) {return (as.vector(cleaned_nodes$adoption_date ==  month))}
  else {return (as.vector(cleaned_nodes$adoption_date > month | cleaned_nodes$adoption_date == "Inf"))}
}


as.vector(cleaned_nodes$adoption_date > 8 | cleaned_nodes$adoption_date == "Inf") #gives a vector of all of the doctors who started prescribing after month 8 or who never started prescribing 




```

8b. Compute the average proportions from (8a) for each month in the study. Use an apply function rather than a loop if you can. Plot the two average proportions from (8a) over time, and in a second plot show their difference. Do the doctors who adopt in a given month consistently have more contacts who are already prescribing than the non-adopters?

```{r}
all.month.prop = apply(as.array(1:17), 1, averaging.peer.pressure) ##apply the averaging.peer.pressure function for every month from 1 to 17

samemonthprop = function(month){
mean(apply(as.array(which(cleaned_nodes$adoption_date == month)), 1, prop_peer_pressure), na.rm = TRUE) 
}

latermonthsprop = function(month){
  mean(apply(as.array(which(cleaned_nodes$adoption_date > month |  cleaned_nodes$adoption_date == "Inf")), 1, prop_peer_pressure), na.rm = TRUE)
}

all.month.same = apply(as.array(1:17), 1, samemonthprop)
all.month.same

all.month.later = apply(as.array(1:17), 1, latermonthsprop)
all.month.later

y1 = all.month.same
y2 = all.month.later

##one way to plot 
plot(x = c(1:17), y1, ylim=range(c(y1,y2)), col = "red", xlab = "month", ylab = "proportion of contacts prescribing")
par(new = TRUE)
plot(x = c(1:17), y2, ylim = range(c(y1,y2)), col = "green", xlab = "month", ylab = "proportion of contacts prescribing")



##alternate way to plot
?data.frame

d = data.frame(x = c(1:17), y1=all.month.same, y2=all.month.later)
library(ggplot2)
ggplot(d, aes(x=x, y=value, color=variable)) + 
geom_point(aes(y=y1, col="y1", pch=2 ))+
geom_point(aes(y=y2, col="y2", pch=1)) + scale_shape_identity() + labs(list(x="month", y="proportion of contacts who have already started prescribing"))
```

```{r}
diff = all.month.same - all.month.later
diff

plot(diff)
```

